Modeling Macroeconomic Phenomena \\with Multi-Agent Reinforcement Learning

### 2019-12-06 paper/paper.tex
Fixed policy gradient section with explanation

@@

\subsubsection{REINFORCE Policy Gradient}
Perhaps the simplest policy gradient method is the REINFORCE algorithm. This algorithm uses a Monte Carlo approach to collecting training data. The length of the episode $T$ is manually specified, and we generally used $T \sim 100$.
\begin{algorithm}
\caption{REINFORCE Policy Gradient \cite{reinforce-lectures}}\label{euclid}
\begin{algorithmic}[1]
\Procedure{REINFORCE}{}
\State $\text{Initialize policy network parameters }\theta$
\For {$i = 1, 2, \ldots, M$}
\State $\text{Follow current policy to obtain trajectory }$\newline\phantom{======}$\{s_1, a_1, r_1, \ldots, s_T, a_T, r_T\}$.
\For {$t = 1, 2, \ldots, T$}
\State $\text{Compute }v_t = \sum_{i = t}^n \gamma^{i- t} r_i$.
\State $\text{Update } \theta \leftarrow \theta + \alpha \nabla_{\theta}\log\pi_{\theta} (s_t, a_t) v_t$
\EndFor
\EndFor
\State \Return {$\theta$}
\EndProcedure
\end{algorithmic}
\end{algorithm}
The runtime of this algorithm per episode is relatively short. However, often a significant number of episodes are required for the following reasons:
\begin{itemize}
  \item Agents only update the parameters once per episode.
  \item In many contexts, vanilla REINFORCE has been known to suffer from high variance in the losses from episode to episode.\cite{reinforce-lectures}
  \item In our multi-agent problem, the environment dynamics constantly change as every agent responds to each timestep in the markets.
\end{itemize}

@@

Recall that the loss function for the REINFORCE algorithm is of the following form:
\[
  \mathcal{L} = -\sum_{t = 1}^T \left[\log \pi_\theta(s_t, a_t) \left( \sum_{i = t}^n \gamma^{i - t}r_i \right) \right].  
\]

In particular, as the policy converges, the probability of taking a given action $a_t$ from state $s_t$ should approach one, and so the $\log \pi_\theta$ term will approach zero. Therefore, convergence corresponds to the losses approaching zero. This occurs in our preliminary simulation, as can be seen in figure \ref{plt:losstozero}.

\begin{figure}[h]
\begin{tikzpicture}[every mark/.append style={mark size=1pt}]
  \begin{axis}[
    xlabel=Episode,
    ylabel=Loss,
    legend style={at={(1,0)}, anchor=south east}
  ]
  \addplot table [y=fl, x=i]{prelims.dat};
  \addlegendentry{Firm}
  \addplot table [y=pl, x=i]{prelims.dat};
  \addlegendentry{Human (Mean)}
  \end{axis}
\end{tikzpicture}
\caption{The agents' losses converge to zero over time. This plot illustrates the loss of firm agents vs. human agents in a simulation with one firm agent and five human agents. The loss for the human agents is the average of all five human agents' losses. This simulation uses the REINFORCE learning algorithm, and indicates that the algorithm is succcessfully converging to a policy within 100 episodes.}
\label{plt:losstozero}
\end{figure}

>>>>

In policy gradient methods, instead of approximating the value functions and using a separate exploration strategy for choosing actions, we directly compute the policy. Given a policy model parameterized by $\theta$, we compute probabilities $\pi_{\theta}(a|s)$. When training this model, we aim to minimize the loss function \cite{actor-critic}:
\[
J(\theta) = -\sum_{s \in S} P_{\pi}(s) \sum_{a\in A} \pi_{\theta}(a|s) Q^{\pi}(s, a)
\]
Intuitively, a low loss means that the agent outputs high probabilities for high $Q$-valued actions. Policy Gradient Theorem allows us to express the gradient of this loss in a tractable manner \cite{reinforce-lectures} \cite{actor-critic}:
\[
\nabla_{\theta} J(\theta) = \mathbb{E}_{\pi}[Q^{\pi}(s, a) \nabla_{\theta} \log \pi_{\theta}(a|s)]
\]
This means that over a trajectory, we simply need to compute the gradients:
\[
\left(\sum_{i=t}^T \gamma^{i-t} r_t\right) \nabla_{\theta} \log \pi_{\theta}(a_t|s_t)
\]
We can then use standard gradient descent methods to optimize $\theta$.

@@

\subsubsection{REINFORCE Policy Gradient}
Perhaps the simplest policy gradient method is the REINFORCE algorithm. This algorithm uses a Monte Carlo approach to collecting training data. The length of the episode $T$ is manually specified, and we generally used $T \sim 100$.
\begin{algorithm}
\caption{REINFORCE Policy Gradient \cite{reinforce-lectures}}\label{euclid}
\begin{algorithmic}[1]
\Procedure{REINFORCE}{}
\State $\text{Initialize policy network parameters }\theta$
\For {$i = 1, 2, \ldots, M$}
\State $\text{Follow current policy to obtain trajectory }$\newline\phantom{======}$\{s_1, a_1, r_1, \ldots, s_T, a_T, r_T\}$.
\For {$t = 1, 2, \ldots, T$}
\State $\text{Compute }v_t = \sum_{i = t}^T \gamma^{i- t} r_i$.
\State $\text{Update } \theta \leftarrow \theta + \alpha \nabla_{\theta}\log\pi_{\theta} (s_t, a_t) v_t$
\EndFor
\EndFor
\State \Return {$\theta$}
\EndProcedure
\end{algorithmic}
\end{algorithm}
The runtime of this algorithm per episode is relatively short. However, often a significant number of episodes are required for the following reasons:
\begin{itemize}
  \item Agents only update the parameters once per episode.
  \item In many contexts, vanilla REINFORCE has been known to suffer from high variance in the losses from episode to episode.\cite{reinforce-lectures}
  \item In our multi-agent problem, the environment dynamics constantly change as every agent responds to each timestep in the markets.
\end{itemize}

@@

Recall that the gradient the REINFORCE algorithm is of the following form:
\[
  \nabla_{\theta} J(\theta) = -\sum_{t = 1}^T \left[\nabla_{\theta} \log \pi_\theta(s_t, a_t) \left( \sum_{i = t}^n \gamma^{i - t}r_i \right) \right].  
\]

In our training, we designate a proxy for the loss to reflect this gradient:
\[
  J'(\theta) = -\sum_{t = 1}^T \left[\log \pi_\theta(s_t, a_t) \left( \sum_{i = t}^n \gamma^{i - t}r_i \right) \right].  
\]
Note that, as the policy converges, the probability of taking a given action $a_t$ from state $s_t$ should approach one, and so the $\log \pi_\theta$ term will approach zero. Therefore, convergence corresponds to the (proxy) losses approaching zero. This occurs in our preliminary simulation, as can be seen in figure \ref{plt:losstozero}.

\begin{figure}[h]
\begin{tikzpicture}[every mark/.append style={mark size=1pt}]
  \begin{axis}[
    xlabel=Episode,
    ylabel=Loss (Proxy),
    legend style={at={(1,0)}, anchor=south east}
  ]
  \addplot table [y=fl, x=i]{prelims.dat};
  \addlegendentry{Firm}
  \addplot table [y=pl, x=i]{prelims.dat};
  \addlegendentry{Human (Mean)}
  \end{axis}
\end{tikzpicture}
\caption{The agents' losses converge to zero over time. This plot illustrates the loss of firm agents vs. human agents in a simulation with one firm agent and five human agents. The loss for the human agents is the average of all five human agents' losses. This simulation uses the REINFORCE learning algorithm, and indicates that the algorithm is succcessfully converging to a policy within 100 episodes.}
\label{plt:losstozero}
\end{figure}
